---
title: "Assignment 4_Species Distribution Model"
author: "Allison White"
date: "January 31, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Objectives
The objective of this analysis is to model the destribution of saguaro cacti (Carnegiea gigantea) across locations with similar environmental conditions.

## Methods
##Species Information
Saguaro cacti are well known tall, almost tree-like cacti which grow in the Sonoran Desert in the southwestern United States. Like other cacti, saguaros have deep taproots and can store large amounts of water in order to survive hot, arid climates.

##Statistical Analysis
Saguaro distribution was modelled using the BIOCLIM method. This modelling approach uses locations of known occurences of a species to identify locations with similar environmental conditions. In this way, locations are weighted based on their suitability for the survival of the species in question. 

```{r,echo=FALSE,include=FALSE}
library('dismo')
library('maptools')
library('rgdal')
library('raster')
library('sp')

bioclim.data <- getData(name = "worldclim",
                        var = "bio",
                        res = 2.5,
                        path = "C:/Users/allwhite/Desktop/Quantitative Ecology Workshop/Species Distribution Models/Data/")

# Read in saguaro observations
obs.data <- read.csv(file = "C:/Users/allwhite/Desktop/Quantitative Ecology Workshop/Species Distribution Models/Data/SDM_Data.csv")

# Notice NAs - drop them before proceeding
obs.data <- obs.data[!is.na(obs.data$latitude), ]

# Determine geographic extent of our data
max.lat <- ceiling(max(obs.data$latitude))
min.lat <- floor(min(obs.data$latitude))
max.lon <- ceiling(max(obs.data$longitude))
min.lon <- floor(min(obs.data$longitude))
geographic.extent <- extent(x = c(min.lon, max.lon, min.lat, max.lat))

# Load the data to use for our base map
data(wrld_simpl)


# Crop bioclim data to geographic extent of saguaro
bioclim.data <- crop(x = bioclim.data, y = geographic.extent)

# Build species distribution model
try(bc.model <- bioclim(x = bioclim.data, p = obs.data),silent=T)

# Drop unused column
obs.data <- obs.data[, c("latitude", "longitude")]

# Build species distribution model
try(bc.model <- bioclim(x = bioclim.data, p = obs.data),silent=T)

# Reverse order of columns
obs.data <- obs.data[, c("longitude", "latitude")]

# Build species distribution model
try(bc.model <- bioclim(x = bioclim.data, p = obs.data),silent=T)

# Predict presence from model
predict.presence <- dismo::predict(object = bc.model, x = bioclim.data, ext = geographic.extent)


# Use the bioclim data files for sampling resolution
bil.files <- list.files(path = "C:/Users/allwhite/Desktop/Quantitative Ecology Workshop/Species Distribution Models/Data/wc2-5", 
                        pattern = "*.bil$", 
                        full.names = TRUE)

# We only need one file, so use the first one in the list of .bil files
mask <- raster(bil.files[1])

# Randomly sample points (same number as our observed points)
background <- randomPoints(mask = mask,     # Provides resolution of sampling points
                           n = nrow(obs.data),      # Number of random points
                           ext = geographic.extent, # Spatially restricts sampling
                           extf = 1.25)             # Expands sampling a little bit

# Arbitrarily assign group 1 as the testing data group
testing.group <- 1

# Create vector of group memberships
group.presence <- kfold(x = obs.data, k = 5) # kfold is in dismo package

# Separate observations into training and testing groups
presence.train <- obs.data[group.presence != testing.group, ]
presence.test <- obs.data[group.presence == testing.group, ]

# Repeat the process for pseudo-absence points
group.background <- kfold(x = background, k = 5)
background.train <- background[group.background != testing.group, ]
background.test <- background[group.background == testing.group, ]

# Build a model using training data
try(bc.model <- bioclim(x = bioclim.data, p = presence.train),silent=T)

# Predict presence from model (same as previously, but with the update model)
predict.presence <- dismo::predict(object = bc.model, 
                                   x = bioclim.data, 
                                   ext = geographic.extent)

# Use testing data for model evaluation
bc.eval <- evaluate(p = presence.test,   # The presence testing data
                    a = background.test, # The absence testing data
                    model = bc.model,    # The model we are evaluating
                    x = bioclim.data)    # Climatic variables for use by model

# Determine minimum threshold for "presence"
bc.threshold <- threshold(x = bc.eval, stat = "spec_sens")
```
## Results
Suitable locations for saguaro cacti calculated from the bioclim model reveal a potential distribution across parts of Nevada, Arizona, and Mexico (Figure 1). The suitable habitat estimated from this model were dependent on a minimum ocurence probability of the presence of saguaro projected over randomly selected areas with no true presence/absence observation (Table 1).

```{r, echo=FALSE}
predict.presence > bc.threshold

# Plot base map
plot(wrld_simpl, 
     xlim = c(min.lon, max.lon),
     ylim = c(min.lat, max.lat),
     axes = TRUE, 
     col = "grey95")

# Only plot areas where probability of occurrence is greater than the threshold
plot(predict.presence > bc.threshold, 
     add = TRUE, 
     legend = FALSE, 
     col = c(NA, "olivedrab"))

# And add those observations
points(x = obs.data$longitude, 
       y = obs.data$latitude, 
       col = "black",
       pch = "+", 
       cex = 0.75)

# Redraw those country borders
plot(wrld_simpl, add = TRUE, border = "grey5")
box()

legend('topright',c('Observed Saguaro','Suitable Habitat'),pch=c('+',NA),col=c('black','olivedrab'),fill=c(NA,'olivedrab'),border=c(NA,'black'))
mtext('Figure 1: Known occurences of saguaro cacti and their suitable habitat',side=1,line = 2)
mtext('estimated by the bioclim model',side=1,line = 3)


library(gridExtra)
tab<-c('Minimum Threshold',bc.threshold)
par(mfrow=c(1,1),oma=c(1,1,1,1),mai=c(1,1,1,1))
plot(1,1,col='white',xaxt='n',yaxt='n',xlab='',ylab='',bty='n')
grid.table(tab)
text(1,1.2,'Table 1: The minimum occurence probability')
text(1,1.15,'accepted for "presence" of saguaro')
```

## Discussion
The suitable habitat estimated from the bioclim model appears to be well represented by the known observed locations with saguaro cacti. However, as this distribution is highly dependent on the minimum occurence probability which defines whether the environmental conditions are similar enough to those where saguaro are known to grow, more information would be required to make this model robust. This could be achieved by incorporating data points where saguaro cacti were not observed (or absence data). 